\chapter{Pure shift NMR}
\label{chpt:pureshift}

\epigraph{\singlespacing%
Then seating herself with a gentleman on each side, she \ldots{} discussed the possibility of improvements with much animation. Nothing was fixed on---but Henry Crawford was full of ideas and projects, and, generally speaking, whatever he proposed was immediately approved.
}{--- \textsc{Jane Austen}, \textit{Mansfield Park}}

\textit{Pure shift} NMR refers to the technique of acquiring \textit{broadband homodecoupled} NMR spectra, where all multiplet structure is suppressed and each chemical environment gives rise to a singlet.\autocite{Zangger2015PNMRS,Castanar2017MRC}
Here, `broadband' means that the couplings are removed from the entire spectrum, and `homodecoupled' refers to the removal of homonuclear couplings: unlike heteronuclear couplings, these cannot be removed by applying RF pulses during the acquisition period, as that will destroy the desired signal itself.
Finally, although pure shift techniques can be applied to any nuclide, \proton{} spectra are of greatest interest because of the narrow chemical shift range of \proton{} which often leads to peak overlap, as well as the abundance of \proton{}--\proton{} couplings in typical organic molecules.

In the first two sections, I first cover the theory underpinning, and a brief history of, pure shift experiments.
I then describe my work towards increasing the quality of pure shift experiments: this is measured both in terms of \textit{sensitivity} as well as \textit{purity}, i.e.\ the lack of spectral artefacts arising from imperfect decoupling.
Each of the approaches described here is compared against the PSYCHE pure shift method, which is the current state of the art.
Finally, I end with a section discussing the combination of pure shift diffusion spectroscopy---formally a pseudo-3D experiment---with the use of ultrafast NMR techniques to collapse the diffusion dimension.
This last project was carried out in collaboration with \JND{} (University of Nantes).

The work in this chapter has not been submitted for publication.

\clearpage

\input{pureshift/intro_theory}
\input{pureshift/intro_practice}
\input{pureshift/nsaltire}
\input{pureshift/optimisation}
\input{pureshift/timerev}
\input{pureshift/dpsyche}
\input{pureshift/epsidosy}

\printbibliography[heading=subbibnumbered]{}
