\chapter{NOAH}

This final---but long---chapter describes my work on \textit{NOAH} (NMR by Ordered Acquisition using \proton{} detection) \textit{supersequences}, pulse sequences which record multiple 2D datasets in the time required for one.\autocite{Kupce2017ACIE}
This is an attractive NMR technique for several reasons: the time savings (which also translate into sensitivity per unit time increases) are clearly a key factor, but the flexibility of combining almost any set of 2D experiments (`modules') also makes NOAH supersequences applicable to a variety of contexts.
Furthermore, the data obtained after processing is virtually identical to well-known standard 2D experiments, meaning that the results are intuitive even to non-experts in NMR.

I begin by introducing the concepts underlying NOAH supersequences, as well as a general discussion of the time savings (and sensitivity per unit time) benefits thus realised.
I then describe the GENESIS (GENEration of Supersequences In Silico) website, which allows users to trivially generate NOAH supersequences containing arbitrary modules.
After this, my work on various aspects of NOAH supersequences is described, with a special focus on newly developed and/or improved modules.
Finally, the design of `parallel' supersequences which use interleaved and/or time-shared modules is discussed.

This work was done in close collaboration with {\=E}riks Kup{\v{c}}e (Bruker UK).
However, all results and analysis shown in this thesis are mine, unless explicitly stated.
The work in this chapter forms the subject of several publications:

\begin{itemize}
    \item \fullcite{Yong2021JMR}
    \item \fullcite{Kupce2021JACSA}
    \item \fullcite{Yong2022AC}
    \item ABBS paper to be submitted
\end{itemize}

The material in the introductory sections also closely follow two reviews which I have contributed to:
\begin{itemize}
    \item \fullcite{Kupce2021NRMP}
    \item RSC book chapter
\end{itemize}

\input{noah/introduction.tex}
\input{noah/snr.tex}
\input{noah/genesis.tex}
\input{noah/modules.tex}
\input{noah/solvsupp.tex}
\input{noah/shortd1.tex}
\input{noah/parallel.tex}

\printbibliography[heading=subbibnumbered]{}
